---
title: "Mining the C++ Standard"
output: 
    html_document:
        toc: true
        toc_float: true
        theme: cosmo
        highlight: tango
---

# Preface

```{r load_libs, echo=FALSE}
library(dplyr, warn.conflicts=FALSE)
library(ggplot2)
library(tidyr)
#library(lubridate, warn.conflicts=FALSE)
library(ggthemes)
library(stringr)

mytheme = theme_bw(base_size=20)##  + theme(
    ## ##text = element_text(family = "Decima WE", color = "grey20"),
    ## ## strip.background = element_blank(),
    ## ## strip.text = element_text(hjust = 0),
    ## ## panel.grid.major = element_line(colour="grey50",size=0.35),
    ## ## panel.grid.minor = element_blank(),
    ## ## plot.margin=unit(c(0,1,0,1),"cm"),
    ## legend.position="top",
    ## plot.caption=element_text(hjust=1,size=14,colour="grey30"),
    ## plot.subtitle=element_text(face="italic",size=14,colour="grey40"),
    ## plot.title=element_text(size=18,face="bold")
## )
df = read.csv("../extracting_impldef/std.csv") %>%
  separate(content_sec0,c("prefix_sec0","stem_sec0",NA,"title_sec0",NA),sep="[\\[\\]\\{\\}]+", remove=F) %>%
  separate(content_sec1,c("prefix_sec1","stem_sec1",NA,"title_sec1",NA),sep="[\\[\\]\\{\\}]+", remove=F) %>%
  separate(content_sec2,c("prefix_sec2","stem_sec2",NA,"title_sec2",NA),sep="[\\[\\]\\{\\}]+", remove=F) %>%
  separate(content_sec3,c("prefix_sec3","stem_sec3",NA,"title_sec3",NA),sep="[\\[\\]\\{\\}]+", remove=F) 
  

levels123 = df  %>%
  select(location_sec0,stem_sec0,
         location_sec1,stem_sec1,
         location_sec2,stem_sec2,
         location_sec3,stem_sec3,
         location_impldef,content_impldef) %>%
  separate(location_impldef,c("file_impldef","lineno_impldef"),sep=":")


## selecting only parts of the standard related to stdlib
stdlib_chapters = c("lib-intro", "support", "diagnostics", "utilities", "strings", "locales", "containers", "iterators", "algorithms", "numerics", "iostreams", "regex", "atomics", "threads", "grammar", "limits", "compatibility", "future")

stdlib_files = c("lib-intro.tex", "support.tex", "diagnostics.tex", "utilities.tex", "strings.tex", "locales.tex", "containers.tex", "iterators.tex", "algorithms.tex", "numerics.tex", "iostreams.tex", "regex.tex", "atomics.tex", "threads.tex", "grammar.tex", "limits.tex", "compatibility.tex", "future.tex")

cat("stdlib files:\n",paste(stdlib_files))

stdlib_123 = levels123 %>% filter(file_impldef %in% stdlib_files)
glimpse(stdlib_123)
```

# Count by tex file

For now, let's start by counting the number of impldefs per chapter!


``` {r num_per_file, echo=FALSE, fig.width=10}


n_plot = ggplot(stdlib_123,aes(file_impldef)) + 
geom_bar() +
xlab("stdlib chapter") +
ylab("N(impldef)") +
coord_flip()
n_plot

num_per_file = stdlib_123 %>% 
    group_by(file_impldef) %>% 
    summarize(count = n()) %>% 
    arrange(count) %>% 
    mutate(fname = factor(file_impldef, file_impldef))


sn_plot = ggplot(num_per_file,aes(x=fname,y=count)) + 
             geom_col() +
             xlab("C++17 STD library file") +
             ylab("N (impldef)") +
             coord_flip() +
             geom_text(aes(label = count, y = count+.5), hjust = "left", size =6) +
             theme(text = element_text(size=20)) 

ggsave("impldef_by_texfile.svg",sn_plot,width=10)
sn_plot

```

# Count by chapter stem

Next, let's find the section stem that an impldef is contained in.

``` {r prim_section_stem, echo=FALSE, fig.width=10, fig.heigth=7}

#obtained with 
#egrep 'rSec1' lib-intro.tex support.tex diagnostics.tex utilities.tex strings.tex locales.tex containers.tex iterators.tex algorithms.tex numerics.tex iostreams.tex regex.tex atomics.tex threads.tex grammar.tex limits.tex compatibility.tex future.tex|sed -e 's/.*\[\(.*\)\].*/"\1", /' | tr -d '\n'

all_sec1_stems = c("library.general", "library.c", "definitions", "description", "requirements", "support.general", "support.types", "support.limits", "cstdint", "support.start.term", "support.dynamic", "support.rtti", "support.exception", "support.initlist", "support.runtime", "diagnostics.general", "std.exceptions", "assertions", "errno", "syserr", "utilities.general", "utility", "intseq", "pairs", "tuple", "optional", "variant", "any", "bitset", "memory", "smartptr", "mem.res", "allocator.adaptor", "function.objects", "meta", "ratio", "time", "type.index", "execpol", "strings.general", "char.traits", "string.classes", "string.view", "c.strings", "localization.general", "locale.syn", "locales", "locale.categories", "c.locales", "containers.general", "container.requirements", "sequences", "associative", "unord", "container.adaptors", "iterators.general", "iterator.requirements", "iterator.synopsis", "iterator.primitives", "predef.iterators", "stream.iterators", "iterator.range", "iterator.container", "algorithms.general", "algorithm.syn", "algorithms.requirements", "algorithms.parallel", "alg.nonmodifying", "alg.modifying.operations", "alg.sorting", "alg.c.library", "numerics.general", "numerics.defns", "numeric.requirements", "cfenv", "complex.numbers", "rand", "numarray", "numeric.ops", "c.math", "input.output.general", "iostreams.requirements", "iostream.forward", "iostream.objects", "iostreams.base", "stream.buffers", "iostream.format", "string.streams", "file.streams", "filesystems", "c.files", "re.general", "re.def", "re.req", "re.syn", "re.const", "re.badexp", "re.traits", "re.regex", "re.submatch", "re.results", "re.alg", "re.iter", "re.grammar", "atomics.general", "atomics.syn", "atomics.alias", "atomics.order", "atomics.lockfree", "atomics.types.generic", "atomics.nonmembers", "atomics.flag", "atomics.fences", "thread.general", "thread.req", "thread.threads", "thread.mutex", "thread.condition", "futures", "gram.key", "diff.iso", "diff.cpp03", "diff.cpp11", "diff.cpp14", "diff.library", "depr.static_constexpr", "depr.impldec", "depr.except.spec", "depr.cpp.headers", "depr.c.headers", "depr.str.strstreams", "depr.uncaught", "depr.func.adaptor.binding", "depr.default.allocator", "depr.storage.iterator", "depr.temporary.buffer", "depr.meta.types", "depr.iterator.primitives", "depr.util.smartptr.shared.obs", "depr.locale.stdcvt", "depr.conversions")

n_sec1_stem = stdlib_123 %>% 
    group_by(stem_sec1) %>% 
    summarize(count = n()) %>% 
    arrange(count)

all_df = data.frame(stem_sec1=all_sec1_stems,count=0) %>% filter(!(stem_sec1 %in% n_sec1_stem$stem_sec1))

glimpse(n_sec1_stem)
glimpse(all_df)


sn_plot = n_sec1_stem %>% 
    mutate(sname = factor(stem_sec1, stem_sec1)) %>%
    ggplot(aes(x=sname,y=count)) + 
             geom_col() +
             xlab("Chapter Stem") +
             ylab("N (impldef)") +
             coord_flip() +
             geom_text(aes(label = count, y = count+.5), hjust = "left", size =4) +
             theme(text = element_text(size=16)) 

ggsave("impldef_by_sec1_stem.svg",sn_plot,width=10)
sn_plot

nrow(n_sec1_stem)
nrow(n_sec1_stem %>% filter(stem_sec1 %in% levels(all_df$stem_sec1)))

all_plot = n_sec1_stem %>% 
            rbind(all_df) %>%
            arrange(count) %>%
            mutate(sname = factor(stem_sec1, stem_sec1)) %>%
            ggplot(aes(x=sname,y=count)) + 
             geom_col() +
             xlab("Chapter Stem") +
             ylab("N (impldef)") +
             coord_flip() +
             #geom_text(aes(label = count, y = count+.5), hjust = "left", size =4) +
             theme(text = element_text(size=16)) 

ggsave("impldef_by_sec1_stem.svg",all_plot,width=10)
all_plot

```

# Selected chapters

## unord

``` {r unord, echo=FALSE, fig.width=10, fig.heigth=7}

hilitd = n_sec1_stem %>% 
    mutate(sname = factor(stem_sec1, stem_sec1), highlight_this = ifelse(stem_sec1 == "unord", 1, 0)) %>%
    ggplot(aes(x=sname,y=count,fill=highlight_this)) + 
             geom_col() +
             xlab("Chapter Stem") +
             ylab("N (impldef)") +
             coord_flip() +
             geom_text(aes(label = count, y = count+.5), hjust = "left", size =4) +
             theme(text = element_text(size=16)) +
             theme(legend.position="none")
hilitd

ggsave("impldef_by_sec1_stem_unord_hilitd.svg",all_plot,width=10)

uodf = stdlib_123 %>% 
    filter( str_detect(as.character(stem_sec1),"unord") )
    
unique(uodf$stem_sec1)
unique(uodf$stem_sec2)
```

Checking the invidual impldef statements in `containers.tex` reveals, that these statements are of the type:

> `For the default constructor, the number of buckets is \impldef{default number of buckets in \tcode{unordered_set}}.`

## char.traits

``` {r chartraits, echo=FALSE, fig.width=10, fig.heigth=7}

hilitd = n_sec1_stem %>% 
    mutate(sname = factor(stem_sec1, stem_sec1), highlight_this = ifelse(stem_sec1 == "char.traits", 1, 0)) %>%
    ggplot(aes(x=sname,y=count,fill=highlight_this)) + 
             geom_col() +
             xlab("Chapter Stem") +
             ylab("N (impldef)") +
             coord_flip() +
             geom_text(aes(label = count, y = count+.5), hjust = "left", size =4) +
             theme(text = element_text(size=16)) +
             theme(legend.position="none")
hilitd
ggsave("impldef_by_sec1_stem_chartraits_hilitd.svg",all_plot,width=10)

ctdf = stdlib_123 %>% 
    filter( str_detect(as.character(stem_sec1),"char.traits") )
    
unique(ctdf$stem_sec1)
unique(ctdf$stem_sec2)
unique(ctdf$file_impldef)

```

So checking `strings.tex` is the next step. The impldef statements here revolve around [char.traits](https://en.cppreference.com/w/cpp/string/char_traits) and how types used by these traits are defined or how Unicode can be handled.

## c.math

``` {r cmath, echo=FALSE, fig.width=10, fig.heigth=7}

hilitd = n_sec1_stem %>% 
    mutate(sname = factor(stem_sec1, stem_sec1), highlight_this = ifelse(stem_sec1 == "c.math", 1, 0)) %>%
    ggplot(aes(x=sname,y=count,fill=highlight_this)) + 
             geom_col() +
             xlab("Chapter Stem") +
             ylab("N (impldef)") +
             coord_flip() +
             geom_text(aes(label = count, y = count+.5), hjust = "left", size =4) +
             theme(text = element_text(size=16)) +
             theme(legend.position="none")
hilitd
ggsave("impldef_by_sec1_stem_cmath_hilitd.svg",all_plot,width=10)

cmdf = stdlib_123 %>% 
    filter( str_detect(as.character(stem_sec1),"c.math") )
    
unique(cmdf$stem_sec1)
unique(cmdf$stem_sec2)
unique(cmdf$file_impldef)
cmdf[,-3:-1]
```

All of these impldef statements are contained in `numerics.tex`. Most of them revolve around [specialized math functions](https://en.cppreference.com/w/cpp/numeric/special_math) and how they handle ill-posed inputs, e.g.

> `\impldef{effect of calling associated Legendre polynomials with \\tcode{l >= 128}}`

## rand

``` {r rand, echo=FALSE, fig.width=10, fig.heigth=7}

hilitd = n_sec1_stem %>% 
    mutate(sname = factor(stem_sec1, stem_sec1), highlight_this = ifelse(stem_sec1 == "rand", 1, 0)) %>%
    ggplot(aes(x=sname,y=count,fill=highlight_this)) + 
             geom_col() +
             xlab("Chapter Stem") +
             ylab("N (impldef)") +
             coord_flip() +
             geom_text(aes(label = count, y = count+.5), hjust = "left", size =4) +
             theme(text = element_text(size=16)) +
             theme(legend.position="none")
hilitd
ggsave("impldef_by_sec1_stem_cmath_hilitd.svg",all_plot,width=10)

rndf = stdlib_123 %>% 
    filter( str_detect(as.character(stem_sec1),"rand") )
    
unique(rndf$stem_sec1)
unique(rndf$stem_sec2)
unique(rndf$file_impldef)
rndf[,-3:-1]
```

All of these impldef statements are contained in `numerics.tex`.

- type of default `random_device` is impldef (so better always specify)
- exception type when `random_device` fails to construct or doesn't work
- how `random_device::operator()` generates values (thread-safety?)

## time

``` {r time, echo=FALSE, fig.width=10, fig.heigth=7}

hilitd = n_sec1_stem %>% 
    mutate(sname = factor(stem_sec1, stem_sec1), highlight_this = ifelse(stem_sec1 == "time", 1, 0)) %>%
    ggplot(aes(x=sname,y=count,fill=highlight_this)) + 
             geom_col() +
             xlab("Chapter Stem") +
             ylab("N (impldef)") +
             coord_flip() +
             geom_text(aes(label = count, y = count+.5), hjust = "left", size =4) +
             theme(text = element_text(size=16)) +
             theme(legend.position="none")
hilitd
ggsave("impldef_by_sec1_stem_cmath_hilitd.svg",all_plot,width=10)

tdf = stdlib_123 %>% 
    filter( str_detect(as.character(stem_sec1),"time") )
    
unique(tdf$stem_sec1)
unique(tdf$stem_sec2)
unique(tdf$file_impldef)
tdf[,-3:-1]
```

All 2 hits are inside `time.tex`. **C-style times** considered harmful!

> `It is \\impldef{whether values are rounded or truncated to the required precision when converting between \\tcode{time_t}`

## Misc finds

- The return value of `std::pow(0.,0.)` is impldef! (numerics.tex:1199)

- often find exception types are impldef

``` {r exctype, echo=FALSE, fig.width=10, fig.heigth=7}

etdf = stdlib_123 %>% 
    filter( str_detect(content_impldef,"exception"))
    
glimpse(etdf)
etdf[,-3:-1]
```
